---
title: "ESM 206 lab 5"
author: "Yani Pohl"
date: "10/29/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#### Lab 5 objecttives

- learn to parse dates with 'lubridate'
- gets counts of obersvations with count()
- uncount() rows with uncount
- 1 and 2 sample t test
- create a heat map with geom_tile()

```{r, include = FALSE}

# Read in data and attach packages
library(tidyverse)
library(here)
library(janitor)

lobster_abundance <- read_csv(here("data", "lobster_abundance.csv"), 
                              na = "-99999") %>% 
  clean_names()

```


Use the tidyr uncount() function to convert our loberster data from frequency format to case format

```{r}

lobster_tidy <- lobster_abundance %>% 
  uncount(lobster_count)
```


##### Exploratory data visulation

We are only going to consider "site" as out variable of interest

```{r}


ggplot(lobster_tidy, aes(x = site, y = size_mm)) +
  geom_jitter(aes(color = site),
              alpha = 0.5,
              width = 0.2)

# Histogram

ggplot(lobster_tidy, aes(x = size_mm)) +
  geom_histogram(aes(fill = site)) +
  facet_wrap(~site, scales = "free")

ggplot(lobster_tidy, aes(sample = size_mm)) +
  geom_qq() +
  facet_wrap(~site)
```


#### Convert the 'date' colu,m to class 'data'

Use lubridate package to convert to date format and then to help us easily parse month & year

```{r}

lobster_date <- lobster_tidy %>% 
  mutate(
    date_new = lubridate::mdy(date)
  )
```


Now let's parse year and month using lubriate:: month() and lubridate:: year()

```{r}

lobster_parse_date <- lobster_date %>% 
  mutate(
    obs_month = lubridate::month(date_new, label = TRUE),
    obs_year = lubridate::year(date_new)
  )
```

Now lets find counts of observed lobsters based on different hierarchical groupings:

Frist lets count lobsters by year and month
```{r}

lobster_ym <- lobster_parse_date %>% 
  count(obs_year, obs_month)

lobster_y <- lobster_parse_date %>% 
  count(obs_year)

lobster_site <- lobster_parse_date %>% 
  count(site)

```

If we want to create a cummery table that caontains statistics other than counts by groups, it's easier to use group_by() + n()

```{r}

lobster_summary <- lobster_parse_date %>% 
  group_by(site) %>% 
  summarize(
    mean_size = mean(size_mm, na.rm = TRUE),
    sd_size = sd(size_mm, na.rm = TRUE),
    lobster_number = n()
  )

```


#### Find confidence intervals

Use t.test() function to find confidence intervals (for one smaple) and perform t-teshs to compare means of two samples

```{r}

ivee_lobster <- lobster_tidy %>% 
  filter(site == "IVEE") %>% 
  pull(size_mm)

t.test(ivee_lobster)
```


#### Two Sample T-Test to Compare Means

Is there a signfigant difference n lobster length at Naples and Mohawk reefs?

```{r}

napl_sample <- lobster_tidy %>% 
  filter(site == "NAPL") %>% 
  pull(size_mm)

mohk_sample <- lobster_tidy %>% 
  filter(site == "MOHK") %>% 
  pull(size_mm)

mn_ttest <- t.test(mohk_sample, napl_sample)
mn_ttest
```

There is a signifigant different in lobster lengths between Naples and Nohawk Reef. (t(df) = statstic, p , .001, alpha = .05)


```{r}

lobster_mn <- lobster_tidy %>% 
  filter(site %in% c("NAPL", "MOHK"))

mn_ttest2 <- t.test(size_mm ~ site, na.rm = TRUE, data = lobster_mn)
mn_ttest2

mn_ttest2$p.value
mn_ttest2$statistic
```

THIS IS CALLED IN LINE REFERANCING - GOOD TOOL TO AUTO- UPDATE FROM CODE

Here is the t-statistic: `r mn_ttest2$p.value`

Here is the t-statistic: `r mn_ttest2$statistic`


#### Now a Heat Map

```{r}

lobster_ys <- lobster_parse_date %>% 
  count(obs_year, site)

ggplot(data = lobster_ys, aes(x = obs_year, y = site)) +
  geom_tile(aes(fill = n))
```

